Appendix

### 2025-04-11 manuscript_rmarkdown/manuscript.tex
Fixed spelling, grammar and style issues.

\title{Location and Scale-Invariant Power Transformations for
Transforming Data to Normality}
\author{Alex Zwanenburg, Steffen Löck}
\date{2025-04-09}

@@

\caption{Effect of location, scale and outliers on estimation of the Box-Cox and Yeo-Johnson transformation parameter $\lambda$. $10000$ samples were drawn from a normal distribution: $\mathcal{N}(\mu, 1)$ for the shift dataset, $\mathcal{N}(10, \sigma)$ for the scale dataset and $\mathcal{N}(0, 1)$ for the outlier dataset. Additionally, an outlier with value $d$ was added to the outlier dataset. Since samples are drawn from a normal distribution, a transformation parameter of $\lambda = 1$ is expected. However, a large shift in location, a scale that is small compared to the location, or presence of large outliers lead to incorrectly estimated transformation parameter values.}\label{fig:decreased-normality}
\end{figure}

@@

\begin{definition}
Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:yeo-johnson-invariant}).
Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed sequence $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.

@@

Here, weighted maximum likelihood estimation is based on equations
\ref{eqn:box-cox-invariant-log-likelihood} and
\ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to Raymaekers
and Rousseeuw (2024), these log-likelihood functions includes additional
terms to accommodate estimation of \(x_0\) and \(s\). The weighted
location- and scale-invariant Box-Cox log-likelihood function is:

\begin{equation}
\label{eqn:box-cox-weighted-invariant-log-likelihood}
\begin{split}
\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
& - \lambda \left( \sum_{i=1}^n w_i \right) \log s + \left( \lambda - 1 \right) \sum_{i=1}^n w_i \log \left[ x_i - x_0 \right]
\end{split}
\end{equation}

where \(\mu_w\) and \(\sigma^2_w\) are the weighted mean and weighted
variance of the Box-Cox transformed feature
\(\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})\):

\begin{equation}
\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
\end{equation}

@@

\begin{equation}
\label{eqn:yeo-johnson-weighted-invariant-log-likelihood}
\begin{split}
\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
& - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
\end{split}
\end{equation}

where \(\mu_w\) and \(\sigma^2_w\) are the weighted mean and weighted
variance of the Yeo-Johnson transformed feature
\(\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})\):

\begin{equation}
\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
\end{equation}

@@

\begin{itemize}
\tightlist
\item
  A step function, with \(\delta_1 \geq 0\) as threshold parameter:
\end{itemize}

\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| \dot{x}_i \right| \leq \delta_1\\
0 & \text{if } \left| \dot{x}_i \right| > \delta_1
\end{cases}
\end{equation}

\begin{itemize}
\tightlist
\item
  A triangle function (or generalised Huber weight), with
  \(\delta_1 \geq 0\) and \(\delta_2 \geq \delta_1\) as threshold
  parameters:
\end{itemize}

\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| \dot{x}_i \right| < \delta_1\\
1 - \frac{\left| \dot{x}_i \right| - \delta_1}{\delta_2 - \delta_1} & \text{if } \delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2 \\
0 & \text{if } \left| \dot{x}_i \right| > \delta_2
\end{cases}
\end{equation}

\begin{itemize}
\tightlist
\item
  A tapered cosine function (J. W. Tukey 1967), with \(\delta_1 \geq 0\)
  and \(\delta_2 \geq \delta_1\) as threshold parameters:
\end{itemize}

\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| \dot{x}_i \right| < \delta_1\\
0.5 + 0.5 \cos\left(\pi \frac{\left| \dot{x}_i \right| - \delta_1}{\delta_2 - \delta_1} \right) & \text{if } \delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2 \\
0 & \text{if } \left| \dot{x}_i \right| > \delta_2
\end{cases}
\end{equation}

@@

Figure \ref{empirical-central-normality-kappa} shows
\(\tau_{\alpha = 0.95, n, \kappa}\) as a function of \(n\) for different
values of \(\kappa\). With decreasing \(\kappa\), the test statistic
curve decreases. The curves for \(\kappa = 0.60\), \(\kappa = 0.70\) and
\(\kappa = 0.80\) are similar, whereas curves for \(\kappa \geq 0.90\)
are affected by outliers, as can be observed by comparing curves of the
central normality test with those of the empirical variant.

@@

To assess type I error rates for the (empirical) central normality test
and compare these to the Shapiro-Wilk test for normality, we randomly
drew another 1000 sequences from a standard normal distribution
(\(\mathcal{N}(0,1)\) for each
\(n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}\).
A second dataset was created by replacing elements with randomly drawn
outliers, as described above. For each sequence, the p-value for the
respective test was used to reject the null hypothesis that the sequence
is (centrally) normal. The null hypothesis was rejected if
\(p \leq 0.05\). The type I error rate was determined by computing the
fraction of centrally normal sequences rejected this way. Figure
\ref{empirical-central-normality-error_rate} shows that the empirical
central normality test had the expected type I error rate for centrally
normal data (with tails of the distribution not being normally
distributed), whereas the central normality test and Shapiro-Wilk test
overestimated type I error rates.

@@

Location- and scale-invariant transformation led to a lower overall
residual errors, indicating a better transformation. Conventional
transformations inflated the mean \(\mu_{YJ}\) and standard deviation
\(\sigma_{YJ}\) of the age feature after transformation. The empirical
central normality test did not detect any statistically significant
deviations from central normality for any transformation (all
\(p \geq 0.93\)).

@@

Considering single process parameters, the choice of learner (Friedman
test: \(p < 10^{-8})\)), normalisation method (Wilcoxon signed rank
test: \(p = 4 \cdot 10^{-8}\)), and transformation method (Friedman
test:\(p = 0.007\), all had a significant impact (at \(p = 0.05\)).

To estimate the marginal effects of process parameters, including
transformation method, we first fit a regression random forest (ranger
package (Wright and Ziegler 2017) version 0.16.0): 2000 trees, node size
2, other hyperparameters default) with process parameters and task
difficulty as predictors and normalised rank as response variable. The
estimated marginal effects are shown in Figure
\ref{fig:marginal-effect-plot}. On the scale of normalised ranks
(\([0.0, 1.0]\)), only the random forest performed better than the
average of all learners (rank difference \(0.173\)). The overall
marginal effect of using z-standardisation to normalise features was
\(0.011\). Transformation methods had the following marginal effects:
\(-0.001\) for using conventional Yeo-Johnson transformation instead of
no transformation; \(-0.012\) for using robust invariant Yeo-Johnson
transformation instead of no transformation; \(-0.012\) for using robust
invariant Yeo-Johnson transformation with empirical central normality
test instead of no transformation; \(-0.012\) for using robust invariant
Yeo-Johnson transformation instead of conventional Yeo-Johnson
transformation; and \(-0.012\) for using using robust invariant
Yeo-Johnson transformation with empirical central normality test instead
of conventional Yeo-Johnson transformation.

@@

Location- and scale-invariant transformations are realised by
simultaneously optimising three parameters, i.e.~transformation
parameter \(\lambda\), shift parameter \(x_0\) and scale parameter
\(s\). We derived the log-likelihood function to facilitate optimisation
using MLE. Alternatively, standardisation of a numeric variable (e.g.,
through subtracting its median value and division by its interquartile
range) prior to conventional power transformations can achieve a similar
effect in reducing sensitivity to the distribution's location and scale,
as demonstrated by our findings. While this alternative helps prevent
these issues -- provided that normalisation does not lead to negative
values for Box-Cox transformations -- location- and scale-invariant
transformations are able to achieve better transformations to normality,
demonstrated by a lower residual error.

@@

We introduced an empirical test for central normality to assess whether
sequences deviate from normality in a way that might require closer
inspection prior to further processing. The empirical test for central
normality differs from other tests for normality, such as the
Shapiro-Wilk test (Shapiro and Wilk 1965), as it assesses normality of
the central portion of a feature, instead of the entire feature.

This work has the following limitations. Firstly, we did observe several
numerical stability issues for optimisation criteria other than MLE
(Appendix B). These appear in regions where transformation parameters
would lead to very large or small numbers when using conventional power
transformations. For MLE stability issues were not observed.

>>>>

\title{Location and Scale-Invariant Power Transformations for
Transforming Data to Normality}
\author{Alex Zwanenburg, Steffen Löck}
\date{2025-04-11}

@@

\caption{Effect of location, scale and outliers on estimation of the Box-Cox and Yeo-Johnson transformation parameter $\lambda$. $10000$ samples were drawn from a normal distribution: $\mathcal{N}(\mu, 1)$ for the \textit{shift} dataset, $\mathcal{N}(10, \sigma)$ for the \textit{scale} dataset and $\mathcal{N}(0, 1)$ for the \textit{outlier} dataset. Additionally, an outlier with value $d$ was added to the outlier dataset. Since samples are drawn from a normal distribution, a transformation parameter of $\lambda = 1$ is expected. However, a large shift in location, a scale that is small compared to the location, or presence of large outliers lead to incorrectly estimated transformation parameter values.}\label{fig:decreased-normality}
\end{figure}

@@

\begin{definition}[Log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation]
Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:yeo-johnson-invariant}).
Let $\mu$ and $\sigma^2$ be the mean and variance of the transformed sequence $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$, respectively.

@@

Here, weighted maximum likelihood estimation is based on equations
\ref{eqn:box-cox-invariant-log-likelihood} and
\ref{eqn:yeo-johnson-invariant-log-likelihood}. Compared to Raymaekers
and Rousseeuw (2024), these log-likelihood functions include additional
terms to accommodate estimation of \(x_0\) and \(s\). The weighted
location- and scale-invariant Box-Cox log-likelihood function is:

\begin{definition}[Weighted log-likelihood function of the location- and scale- invariant Box-Cox power transformation]
Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:box-cox-invariant}).
Let $w_i \geq 0$ be the weight corresponding to each element of $\mathbf{X}$.
Let $\mu_w$ be the weighted mean of the Box-Cox transformed sequence:
\begin{equation*}
\mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
\end{equation*}

Let $\sigma^2_w$ be the weighted variance of the Box-Cox transformed sequence:
\begin{equation*}
\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i}
\end{equation*}

Then, the weighted log-likelihood function of the location- and scale- invariant Box-Cox power transformation is:

\begin{equation}
\label{eqn:box-cox-weighted-invariant-log-likelihood}
\begin{split}
\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
& - \lambda \left( \sum_{i=1}^n w_i \right) \log s + \left( \lambda - 1 \right) \sum_{i=1}^n w_i \log \left[ x_i - x_0 \right]
\end{split}
\end{equation}
\end{definition}

@@

\begin{definition}[Weighted log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation]

Let $\mathbf{X}$, $\lambda$, $x_0$, and $s$ be defined as earlier (Eqn. \ref{eqn:yeo-johnson-invariant}).
Let $w_i \geq 0$ be the weight corresponding to each element of $\mathbf{X}$.
Let $\mu_w$ be the weighted mean of the Yeo-Johnson transformed sequence:
\begin{equation*}
\mu_w = \frac{\sum_{i=1}^n w_i \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
\end{equation*}

Let $\sigma^2_w$ be the weighted variance of the Yeo-Johnson transformed sequence:
\begin{equation*}
\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i}
\end{equation*}

Then, the weighted log-likelihood function of the location- and scale- invariant Yeo-Johnson power transformation is:

\begin{equation}
\label{eqn:yeo-johnson-weighted-invariant-log-likelihood}
\begin{split}
\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
& - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \mathop{\mathrm{sgn}}(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
\end{split}
\end{equation}
\end{definition}

@@

\begin{itemize}
\item
  A step function, with \(\delta_1 \geq 0\) as threshold parameter:
  \begin{equation*}
  w_i =
  \begin{cases}
  1 & \text{if } \left| \dot{x}_i \right| \leq \delta_1\\
  0 & \text{if } \left| \dot{x}_i \right| > \delta_1
  \end{cases}
  \end{equation*}
\item
  A triangle function (or generalised Huber weight), with
  \(\delta_1 \geq 0\) and \(\delta_2 \geq \delta_1\) as threshold
  parameters: \begin{equation*}
  w_i =
  \begin{cases}
  1 & \text{if } \left| \dot{x}_i \right| < \delta_1\\
  1 - \frac{\left| \dot{x}_i \right| - \delta_1}{\delta_2 - \delta_1} & \text{if } \delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2 \\
  0 & \text{if } \left| \dot{x}_i \right| > \delta_2
  \end{cases}
  \end{equation*}
\item
  A tapered cosine function (J. W. Tukey 1967), with \(\delta_1 \geq 0\)
  and \(\delta_2 \geq \delta_1\) as threshold parameters:
  \begin{equation*}
  w_i =
  \begin{cases}
  1 & \text{if } \left| \dot{x}_i \right| < \delta_1\\
  0.5 + 0.5 \cos\left(\pi \frac{\left| \dot{x}_i \right| - \delta_1}{\delta_2 - \delta_1} \right) & \text{if } \delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2 \\
  0 & \text{if } \left| \dot{x}_i \right| > \delta_2
  \end{cases}
  \end{equation*}
\end{itemize}

@@

Figure \ref{fig:empirical-central-normality-kappa} shows
\(\tau_{\alpha = 0.05, n, \kappa}\) as a function of \(n\) for different
values of \(\kappa\). With decreasing \(\kappa\), the test statistic
curve decreases. The curves for \(\kappa = 0.60\), \(\kappa = 0.70\) and
\(\kappa = 0.80\) are similar, whereas curves for \(\kappa \geq 0.90\)
are affected by outliers, as can be observed by comparing curves of the
central normality test with those of the empirical variant.

@@

To assess type I error rates for the (empirical) central normality test
and compare these to the Shapiro-Wilk test for normality, we randomly
drew another 1000 sequences from a standard normal distribution
(\(\mathcal{N}(0,1)\) for each
\(n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}\).
A second dataset was created by replacing elements with randomly drawn
outliers, as described above. For each sequence, the p-value for the
respective test was used to reject the null hypothesis that the sequence
is (centrally) normal. The null hypothesis was rejected if
\(p \leq 0.05\). The type I error rate was determined by computing the
fraction of centrally normal sequences rejected this way. Figure
\ref{fig:empirical-central-normality-error_rate} shows that the
empirical central normality test had the expected type I error rate for
centrally normal data (with tails of the distribution not being normally
distributed), whereas the central normality test and Shapiro-Wilk test
overestimated type I error rates.

@@

Location- and scale-invariant transformation led to a lower overall
residual error, indicating a better transformation to normality.
Conventional transformations inflated the mean \(\mu_{YJ}\) and standard
deviation \(\sigma_{YJ}\) of the age feature after transformation. The
empirical central normality test did not detect any statistically
significant deviations from central normality for any transformation
(all \(p \geq 0.93\)).

@@

Considering single process parameters, the choice of learner (Friedman
test: \(p < 10^{-8})\)), normalisation method (Wilcoxon signed rank
test: \(p = 4 \cdot 10^{-8}\)), and transformation method (Friedman
test: \(p = 0.007\)), all had a significant impact (at \(p = 0.05\)).

To estimate the marginal effects of process parameters, including
transformation method, we first fit a regression random forest (ranger
package (Wright and Ziegler 2017) version 0.16.0): 2000 trees, node size
2, other hyperparameters default) with process parameters and task
difficulty as predictors and normalised rank as response variable. The
estimated marginal effects are shown in Figure
\ref{fig:marginal-effect-plot}. On the scale of normalised ranks
(\([0.0, 1.0]\)), only the random forest performed better than the
average of all learners (rank difference \(0.173\)). The marginal
improvement in performance from z-standardisation was \(0.011\).
Transformation methods had the following marginal effects: \(-0.001\)
for using conventional Yeo-Johnson transformation instead of no
transformation; \(-0.012\) for using robust invariant Yeo-Johnson
transformation instead of no transformation; \(-0.012\) for using robust
invariant Yeo-Johnson transformation with empirical central normality
test instead of no transformation; \(-0.012\) for using robust invariant
Yeo-Johnson transformation instead of conventional Yeo-Johnson
transformation; and \(-0.012\) for using using robust invariant
Yeo-Johnson transformation with empirical central normality test instead
of conventional Yeo-Johnson transformation.

@@

Location- and scale-invariant transformations are realised by
simultaneously optimising three parameters, i.e.~transformation
parameter \(\lambda\), shift parameter \(x_0\) and scale parameter
\(s\). We derived the log-likelihood function to facilitate optimisation
using MLE. Alternatively, standardisation of a numeric variable (e.g.,
through subtracting its median value and division by its interquartile
range) prior to conventional power transformations can achieve a similar
effect in reducing sensitivity to the feature's location and scale.
While this alternative helps prevent these issues -- provided that
standardisation does not lead to negative values for Box-Cox
transformations -- location- and scale-invariant transformations are
able to achieve better transformations to normality, as demonstrated by
lower residual errors in simulation experiments.

@@

We introduced an empirical test for central normality to assess whether
sequences deviate from normality in a way that might require closer
inspection prior to further processing. The empirical test for central
normality differs from other tests for normality, such as the
Shapiro-Wilk test (Shapiro and Wilk 1965), as it assesses normality of
the central portion of a feature, instead of the entire feature.
Compared to the central normality test and the Shapiro-Wilk test, the
empirical central normality test remains consistent in the presence of
outliers, although the former tests are more powerful when outliers are
absent.

This work has the following limitation: We observed several numerical
stability issues for optimisation criteria other than MLE (Appendix B).
These appear in regions where transformation parameters would lead to
very large or small numbers when using conventional power
transformations. For MLE stability issues were not observed.

### 2025-04-11 manuscript_rmarkdown/manuscript_appendix.Rmd
Fixed spelling, grammar and style issues.

The parameters of these power transformations can be optimised based by maximising the log-likelihood function, under the assumption that the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ follows a normal distribution.
The log-likelihood functions for conventional Box-Cox and Yeo-Johnson transformations are well-known.
However, the introduction of scaling parameter $s$ prevents their direct use.
Here, we first derive the general form of the log-likelihood functions, and then derive their power-transformation specific definitions.

@@

Let $\mathbf{X}$ be a feature with ordered feature values, and $\mathbf{Y}^\lambda =\phi^{\lambda} \left(\mathbf{X} \right)$ and $\mathbf{Y}^{\lambda, x_0, s} =\phi^{\lambda, x_0, s} \left(\mathbf{X} \right)$ its transformed values using conventional and shift and scale invariant power transformations, respectively. Since power transformations are monotonic, $\mathbf{Y}$ will likewise be ordered.

Below we will focus on criteria based on the empirical density function and those based on skewness and kurtosis of the transformed featured.
Other potential criteria, such as the Shapiro-Wilk test statistic [@Shapiro1965-zd] are not investigated here.
In the case of the Shapiro-Wilk test statistic this is because of lack of scalability to features with many ($> 5000$) instances, and because adapting the test statistic to include weights is not straightforward.

@@

Here $w_i$ are weights, and $\mu$ and $\sigma$ are location and scale parameters. 
For non-robust power transformations, all $w_i = 1$.
Note that this criterion is not the same as the Anderson-Darling test statistic [@Anderson1952-gz], which involves solving (or approximating) an integral function, contains an extra scalar multiplication term, and does not include weights.
The Anderson-Darling criterion seeks to minimise the squared Euclidean distance between the EDF and the normal CDF, with differences at the upper and lower end of the normal CDF receiving more weight than those at the the centre of the CDF.

@@

Here $w_i$ are weights, and $\mu$ and $\sigma$ are location and scale parameters.
For non-robust power transformations, all $w_i = 1$.
The criterion is similar to the Cramér-von Mises test statistic [@Cramer1928-rc, @Von_Mises1928-ef], aside from a additive scalar value and the introduction of weights.
This criterion, like the Anderson-Darling criterion, seeks to minimise the squared Euclidean distance between the EDF and the normal CDF.
Unlike the Anderson-Darling criterion, this criterion weights all instances equally.

@@

The second class of criteria seeks to reduce skewness and (excess) kurtosis of the transformed feature $\mathbf{Y}$.
We will first define the location $\mu$ and scale $\sigma$ of the the transformed as these are required for computing skewness and kurtosis.
Here, $\mu$ is defined as:

@@

The location, or mean, is weighted using weights $w_i$.
For non-robust transformations, $w_i = 1$.
Then, $\sigma^2$ is defined as:
```{=latex}
\begin{equation}
\sigma^2 = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0, s} \left( x_i \right) - \mu \right)^2}{\sum_{i=1}^n w_i}
\end{equation}
```

@@

\begin{table}
\begin{center}
\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.80}$ for
\textbf{central normality} at $\kappa = 0.80$, as a function of significance 
level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
\label{tab:central-normality-critical-statistic}
\begin{tabular}{l | r r r r r r r r r}

@@

\begin{table}
\begin{center}
\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
\label{tab:empirical-central-normality-critical-statistic}
\begin{tabular}{l | r r r r r r r r r}

@@

The sequences drawn resulted from the permutations of the number of elements of each
sequence ($n \in \{30, 100, 500\}$) and transformation parameter for the
inverse transformation $\lambda \in \{0.1, 1.0, 1.9\}$. Each sequence then underwent
power transformation using Box-Cox and Yeo-Johnson transformations. For Box-Cox
transformations, each sequence was shifted prior to inverse transformation
so that every element after inverse transformation would be strictly positive.

>>>>

The parameters of these power transformations can be optimised by maximising the
log-likelihood function, under the assumption that the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ follows a normal distribution.
The log-likelihood functions for conventional Box-Cox and Yeo-Johnson transformations are well-known.
However, the introduction of scaling parameter $s$ prevents their direct use.
Here, we first derive the general form of the log-likelihood functions, and then derive their power-transformation specific definitions.

@@

Let $\mathbf{X}$ be a feature with ordered feature values,
and $\mathbf{Y}^\lambda =\phi^{\lambda} \left(\mathbf{X} \right)$ and
$\mathbf{Y}^{\lambda, x_0, s} =\phi^{\lambda, x_0, s} \left(\mathbf{X} \right)$ 
its transformed values using conventional and location- and scale-invariant power transformations, respectively.
Since power transformations are monotonic, $\mathbf{Y}$ will likewise be ordered.

Below we will focus on criteria based on the empirical density function and those based on skewness and kurtosis of the transformed feature.
Other potential criteria, such as the Shapiro-Wilk test statistic [@Shapiro1965-zd] are not investigated here.
In the case of the Shapiro-Wilk test statistic this is because of lack of scalability to features with many ($> 5000$) instances, and because adapting the test statistic to include weights is not straightforward.

@@

Here $w_i$ are weights, and $\mu$ and $\sigma$ are location and scale parameters. 
For non-robust power transformations, all $w_i = 1$.
Note that this criterion is not the same as the Anderson-Darling test statistic [@Anderson1952-gz],
which involves solving (or approximating) an integral function,
contains an extra scalar multiplication term, and does not include weights.
The Anderson-Darling criterion seeks to minimise the squared Euclidean distance 
between the EDF and the normal CDF, with differences at the upper and lower end
of the normal CDF receiving more weight than those at the centre of the CDF.

@@

Here $w_i$ are weights, and $\mu$ and $\sigma$ are location and scale parameters.
For non-robust power transformations, all $w_i = 1$.
The criterion is similar to the Cramér-von Mises test statistic [@Cramer1928-rc, @Von_Mises1928-ef],
aside from an additive scalar value and the introduction of weights.
This criterion, like the Anderson-Darling criterion, seeks to minimise the squared Euclidean distance between the EDF and the normal CDF.
Unlike the Anderson-Darling criterion, this criterion weights all instances equally.

@@

The second class of criteria seeks to reduce skewness and (excess) kurtosis of the transformed feature $\mathbf{Y}$.
We will first define the location $\mu$ and scale $\sigma$ of the transformed feature
as these are required for computing skewness and kurtosis.
Here, $\mu$ is defined as:

@@

The location, or mean, is weighted by $w_i$, which equals $1$ for non-robust transformations.

Then, $\sigma^2$ is defined as:
```{=latex}
\begin{equation}
\sigma^2 = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0, s} \left( x_i \right) - \mu \right)^2}{\sum_{i=1}^n w_i}
\end{equation}
```

@@

\begin{table}
\begin{center}
\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.80}$ for
\textbf{central normality} at $\kappa = 0.80$, as a function of significance 
level $\alpha$ and number of instances $n$. The values in this table should be divided by $100$ before use.}
\label{tab:central-normality-critical-statistic}
\begin{tabular}{l | r r r r r r r r r}

@@

\begin{table}
\begin{center}
\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
level $\alpha$ and number of instances $n$.The values in this table should be divided by $100$ before use.}
\label{tab:empirical-central-normality-critical-statistic}
\begin{tabular}{l | r r r r r r r r r}

@@

The sequences drawn resulted from the permutations of the number of elements of each
sequence ($n \in \{30, 100, 500\}$) and transformation parameter for the
inverse transformation $\lambda \in \{0.1, 1.0, 1.9\}$. Each sequence then underwent
power transformation using Box-Cox and Yeo-Johnson transformations. For Box-Cox
transformations, each sequence was shifted prior to inverse transformation
ensuring all elements were strictly positive after inverse transformation.

### 2025-04-11 manuscript_rmarkdown/manuscript_appendix.tex
Fixed spelling, grammar and style issues.

\title{Appendix}
\author{Alex Zwanenburg, Steffen Löck}
\date{2025-04-09}

@@

\section{Appendix A: Log-likelihood functions for location and scale
invariant power
transformation}\label{appendix-a-log-likelihood-functions-for-location-and-scale-invariant-power-transformation}

Location and scale-invariant Box-Cox and Yeo-Johnson transformations are
parametrised using location \(x_0\) and scale \(s\) parameters, in
addition to transformation parameter \(\lambda\). This leads to the
following transformations. The location and scale-invariant Box-Cox
transformation is:

@@

where \(x_i - x_0 > 0\). The location and scale-invariant Yeo-Johnson
transformation is:

>>>>

\title{Appendix}
\author{Alex Zwanenburg, Steffen Löck}
\date{2025-04-11}

@@

\section{Appendix A: Log-likelihood functions for location- and
scale-invariant power
transformation}\label{appendix-a-log-likelihood-functions-for-location--and-scale-invariant-power-transformation}

Location- and scale-invariant Box-Cox and Yeo-Johnson transformations
are parametrised using location \(x_0\) and scale \(s\) parameters, in
addition to transformation parameter \(\lambda\). This leads to the
following transformations. The location- and scale-invariant Box-Cox
transformation is:

@@

where \(x_i - x_0 > 0\). The location- and scale-invariant Yeo-Johnson
transformation is:
